Low Complexity Adaptive Beamformer\\ for Active Sonar Imaging

### 2015-10-13 LCAforActiveSonarImaging/lca.tex
la2pdf commit - 2015:10:13	15:38:04	(Tue13.Oct)

In many cases MVDR can improve the contrast and resolution of active sonar images compared to conventional static methods~\cite{Blomberg2013,Blomberg2012a,Dursun2009,Lo2004}. However, while the computational complexity of conventional beamformers are linear with the number of channels, O($M$), MVDR is at O($M^3$). This makes MVDR impracticable and sometimes infeasible. We have shown in previous studies that MVDR's complexity can be reduced, and its implementation accelerated significantly using graphics computing units (GPUs)~\cite{Buskenes2014,Asen2013}. However, this implementation is complex and still leaves something to be desired in speed.

An alternative to MVDR is the Low Complexity Adaptive (LCA) beamformer. It was first proposed by Synnev\aa{}g in clinical medical imaging~\cite{Synnevag2008}, who demonstrated its ability to obtain very similar results to MVDR in a system with focused transmit beams. LCA applies a set of predefined windows, or tapers, and selects the one offering the best noise suppression. It relies on the same optimization criterion as MVDR, and may be considered a version of MVDR with a reduced window solution space. Alternatively, it may be viewed as a delay-and-sum (DAS) beamformer with an adaptive multi-taper extension. \cite{Stankwitz1995}

>>>>

In many cases MVDR can improve the contrast and resolution of active sonar images compared to conventional static methods~\cite{Blomberg2013,Blomberg2012a,Dursun2009,Lo2004}. However, while the computational complexity of conventional beamformers are linear with the number of channels, O($M$), MVDR is at O($M^3$). This makes MVDR impracticable and sometimes infeasible. We have shown in previous studies that MVDR's complexity can be reduced, and its implementation accelerated significantly using graphics computing units (GPUs)~\cite{Buskenes2014,Asen2013}. However, the implementation is still complex, relatively slow and not intutive.

A light weight alternative to MVDR is the Low Complexity Adaptive (LCA) beamformer. It was first proposed by Synnev\aa{}g in clinical medical imaging~\cite{Synnevag2008}, who demonstrated its ability to obtain very similar results to MVDR in a system with focused transmit beams. LCA applies a set of predefined windows, or tapers, and selects the one offering the best noise suppression. It relies on the same optimization criterion as MVDR, and may be considered a version of MVDR with a reduced window solution space. Alternatively, it may be viewed as a delay-and-sum (DAS) beamformer with an adaptive multi-taper extension. \cite{Stankwitz1995}
